---
title: "EDA of Job Search Data "
author: "Jessica Smith"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(lubridate)
library(knitr)
library(stringr)
library(ggrepel)
library(wordcloud)
library(RColorBrewer)
library(plotly)
library(kableExtra)

set.seed(123)
my_colors <- brewer.pal(10, "Paired") #or dark2
```


```{r include = FALSE, echo = FALSE}
Sys.setenv("plotly_username"="")
Sys.setenv("plotly_api_key"="")
```

## 1. Job Search Activity

```{r read_in_data}

# Load data and ensure columns are correctly formatted
data <- read.csv("job search history.csv")
colnames(data) <- str_trim(colnames(data))
data$applied.date <- as.Date(data$applied.date, format = "%m/%d/%Y")
data$job.post.date <- as.Date(data$job.post.date, format = "%m/%d/%Y")
data$response.date <- as.Date(data$response.date, format = "%m/%d/%Y")

head(data)
```


```{r weekly_applications_plot}

# Add a week_number column by extracting the week number from 'applied date'
data <- data %>%
  mutate(week.number = lubridate::week(applied.date))

# Calculate the minimum date to set as "week 1"
min_week <- min(data$week.number)

# Add relative week number
weekly_applications <- data %>%
  mutate(relative.week.number = week.number - min_week + 1)

# Count total applications and customized resumes per week
weekly_summary <- weekly_applications %>%
  group_by(relative.week.number) %>%
  summarise(
    total_applications = n(),
    customized_resumes = sum(customized.resume. == "Yes", na.rm = TRUE)  # Count customized resumes
  )

# Plot total applications and customized resumes
p1 <- ggplot(weekly_summary, aes(x = relative.week.number)) +
  
  # Line for total applications
  geom_line(aes(y = total_applications, color = "Total Applications"), linewidth = 1) +
  geom_point(aes(y = total_applications, color = "Total Applications",
                 text = paste("Week:", relative.week.number, "<br>Total Applications:", total_applications))) +
  
  # Line for customized resumes
  geom_line(aes(y = customized_resumes, color = "Customized Resumes"), linetype = "dashed", linewidth = 1) +
  geom_point(aes(y = customized_resumes, color = "Customized Resumes", 
                 text = paste("Week:", relative.week.number, "<br>Customized Resumes:", customized_resumes))) +
  
  # Customize the plot
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +  # Ensure whole numbers on x-axis
  labs(title = "Weekly Applications and Customized Resumes", 
       x = "Week Number", 
       y = "Applications",  # Rename Y-axis
       color = "Legend") +  # Legend title
  theme_minimal() +                         # Remove gray background and grid lines
  theme(
    panel.grid.major = element_blank(),     # Remove major grid lines
    panel.grid.minor = element_blank(),     # Remove minor grid lines
    axis.line = element_line(color = "black")  # Add axis lines
  ) +
  scale_color_manual(values = c("Total Applications" = my_colors[1], "Customized Resumes" = my_colors[2])) 


ggsave(p1, 
       filename = "weekly_applications.png",
       device = "png",
       height = 4, width = 6, units = "in")

ggplotly(p1, tooltip = "text")

#api_create(ggplotly(p1, tooltip = "text"), filename = "Weekly Applications")
```



```{r total_applications}

# Count total applications per week
weekly_applications <- data %>%
  mutate(relative.week.number = week.number - min_week + 1) %>%
  group_by(relative.week.number) %>%
  summarise(applications = n())

# Calculate cumulative total number of applications
weekly_applications <- weekly_applications %>%
  arrange(relative.week.number) %>%                       # Sort by week number
  mutate(cumulative_applications = cumsum(applications))  # Calculate cumulative sum

# Define the ratio between the max values of both scales
max_applications <- max(weekly_applications$applications)
max_cumulative <- max(weekly_applications$cumulative_applications)
scale_factor <- max_applications / max_cumulative

# Now plot with two y-axes and labels
p2 <- ggplot(weekly_applications, aes(x = relative.week.number)) +
  geom_line(aes(y = applications), color = my_colors[4]) +  # Line for weekly applications
  geom_point(aes(y = applications), color = my_colors[4]) +  # Points for weekly applications
  
  # Add labels for each data point on the primary y-axis
  geom_text(aes(y = applications, label = applications), vjust = -0.5, color = my_colors[4], size = 3) +

  geom_line(aes(y = cumulative_applications * scale_factor), 
            color = my_colors[2], linetype = "dashed") +  # Line for cumulative applications, scaled
  scale_y_continuous(
    name = "Applications per Week",  # Primary y-axis for weekly applications
    sec.axis = sec_axis(~ . / scale_factor, 
                        name = "Total Applications")  # Secondary y-axis
  ) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +  # Ensure whole numbers on x-axis
  labs(title = "Weekly and Cumulative Applications", x = "Week Number") +
  theme_minimal() +  # Remove gray background and grid lines
  theme(
    panel.grid.major = element_blank(),  # Remove major grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines
    axis.line = element_line(color = "black"),  # Add axis lines
    axis.title.y.left = element_text(color = my_colors[4]),  # Color the primary y-axis label
    axis.text.y.left = element_text(color = my_colors[4]),    # Color the primary y-axis values
    axis.title.y.right = element_text(color = my_colors[2]),  # Color the secondary y-axis label
    axis.text.y.right = element_text(color = my_colors[2])    # Color the secondary y-axis values
  )

p2

ggsave(p2, 
       filename = "weekly_cumulative_applications.png",
       device = "png",
       height = 4, width = 6, units = "in")

```

```{r summary}

# Calculate key metrics
average_applications_per_week <- round(mean(weekly_applications$applications))  # Round to whole number
total_applications <- sum(weekly_applications$applications)
total_interviews <- sum(data$`selected.for.interview.` == "Yes", na.rm = TRUE)
total_number_of_weeks <- max(weekly_applications$relative.week.number)
total_offers = 1
# Create a table using data.frame()
summary_table <- data.frame(
  Metric = c("Weeks Spent Looking for a Job","Total Applications Submitted", "Average Applications per Week", "Applications That Lead to an Interview", "Total Job Offers Recieved"),
  Value = c(total_number_of_weeks, total_applications, average_applications_per_week, total_interviews, total_offers)
)

# Display the table with kable for nice formatting
kable(summary_table, format = "markdown", col.names = c("Metric", "Value"), align = 'l')


```


```{r funnel}

# Create the summary table
summary_table <- data.frame(
  Stage = c("Applications", "Interviews", "Job Offers"),
  Value = c(total_applications, total_interviews, total_offers)
)

# Create a  funnel chart
plot_ly(summary_table, 
        type = "funnel", 
        y = ~Stage,  # Stages on the y-axis
        x = ~Value,  # Values on the x-axis
        textinfo = "value",  # Show both values and percentages
        textposition = "outside",
        marker = list(color = c(my_colors[1], my_colors[2], my_colors[3]))) %>%  # 
  layout(title = "Job Search Outcomes",
         xaxis = list(title = "Count"),
         yaxis = list(title = ""),
         showlegend = FALSE)


```

## 2. Overall Success Rate

```{r overall_response_rate}
library(scales)  # For formatting numbers as percentages

# Calculate response rate (interview requests)
response_rate <- data %>%
  filter(!is.na(`selected.for.interview.`)) %>%
  summarise(success_rate = mean(`selected.for.interview.` == "Yes"))

# Format the success rate as a percentage
formatted_response_rate <- percent(response_rate$success_rate)

# Print the formatted response rate
formatted_response_rate

```

## 3. Customization Performance

### Comparison of response rates

Perform a test to compare the response rates (success rates) for customized resumes versus non-customized resumes. We want to compare whether the response rates (application success) for customized resumes to see if they are significantly different from those for non-customized resumes. 


#### Population Size (Sample Size)
Both groups (customized and non-customized resumes) should have a sufficient number of observations.

```{r population_size}

# Filter out rows where 'customized.resume.' is NA or blank
customization_data <- data[!is.na(data$customized.resume.) & data$customized.resume. != "", ]

# Now count the number of observations in each group
table(customization_data$customized.resume., useNA = "ifany")

```


#### Test for Binary Data
A Chi-squared test would compare the proportions of interviews received for customized vs non-customized resumes. Given that you have only 13 interviews (a small sample for the "Yes" category), Fisher's Exact Test is the better option over the Chi-squared test. This test will help you determine if customizing resumes significantly impacts your likelihood of getting an interview.

Fisher's Exact Test is specifically designed for small sample sizes and is more reliable when dealing with low-frequency data (like only 13 interviews). Unlike the Chi-squared test, Fisher's Exact Test provides an exact p-value rather than an approximation, which is why it's preferred in situations with small expected counts.


```{r contingency_table}

# Create a contingency table of interview results by resume type
contingency_table <- table(customization_data$customized.resume., customization_data$`selected.for.interview.` == "Yes")

contingency_table
```


Use Fisher's Exact Test when one or more of the cells in your contingency table have small expected frequencies (usually less than 5). Since there are only 13 successful applications, Fisher’s Exact Test is more reliable than the Chi-squared test for determining whether there's a significant association between customized resumes and interview success.


```{r fischers_test}
# Perform Fisher's Exact Test
fisher.test(contingency_table)
```

#### Interpretation of Results:
P-value = 1:

A p-value of 1 indicates that there is no statistically significant difference between the interview success rates for customized and non-customized resumes.
In other words, customizing your resume does not significantly increase or decrease the likelihood of getting an interview compared to non-customized resumes, according to this test.
Odds Ratio = 1.072659:

An odds ratio of 1.072659 suggests that the odds of getting an interview with a customized resume are approximately the same as with a non-customized resume.
An odds ratio close to 1 means there is no strong association between customizing the resume and the success rate of getting an interview. Specifically, your odds of success are about 1.07 times higher with a customized resume, but this is not statistically significant.
95% Confidence Interval: [0.1858966, 4.2811602]:

The confidence interval provides a range in which the true odds ratio is likely to fall, with 95% confidence.
In this case, the confidence interval is quite wide, ranging from 0.1859 to 4.2812. This wide range means there is considerable uncertainty about the true effect of customizing resumes.
Crucially, since the interval includes 1, it means the test did not find a significant difference between the two groups. This reinforces the conclusion that there is no strong evidence that customizing resumes affects interview success.

The p-value of 1 indicates that there is no statistically significant association between customizing resumes and getting an interview.
The odds ratio of 1.07 suggests a very small difference in favor of customized resumes, but this is not statistically significant.
The confidence interval is wide and includes 1, further indicating that the difference between customized and non-customized resumes is not meaningful in this case.

Based on the Fisher's Exact Test results, you can conclude that **customizing resumes does not significantly affect the likelihood of getting an interview** in your dataset. There's no strong evidence to suggest that customizing your resume makes a difference in interview success rates.

```{r table}

# Trim spaces from 'customized.resume.'
data$customized.resume. <- str_trim(data$customized.resume.)

# Create the contingency table of interview results by resume type
contingency_table <- table(data$customized.resume., data$`selected.for.interview.` == "Yes")

# Convert contingency table to data frame for easier manipulation
contingency_df <- as.data.frame.matrix(contingency_table)
colnames(contingency_df) <- c("Not_Selected", "Selected")

# Calculate total applications and merge it with the contingency data
contingency_df <- contingency_df %>%
  mutate(`customized.resume.` = rownames(contingency_df),
         total_applications = Not_Selected + Selected) %>%
  select(`customized.resume.`, total_applications, Selected)

# Calculate success rate and merge it with the contingency data
customized_response <- data %>%
  filter(`customized.resume.` != "") %>%
  group_by(`customized.resume.`) %>%
  summarise(success_rate = mean(`selected.for.interview.` == "Yes", na.rm = TRUE))

# Merge contingency data with success rate
final_table <- contingency_df %>%
  left_join(customized_response, by = "customized.resume.") %>%
  mutate(success_rate = round(success_rate * 100, 3))  # Convert success rate to percentage

# Display table with the additional columns
final_table %>%
  kable(col.names = c("Customized?", "Total Applications", "Successful Applications", "Success Rate (%)"),
        caption = "Success of Applications",
        format = "html",  # Switch to HTML format to enable styling
        digits = 4) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE)
```

## 4. Application Responses

```{r no_interview_requests}
# Count applications with no interview request
no_interview_count <- sum(data$`selected.for.interview.` == "No", na.rm = TRUE)
no_interview_count
```

```{r no_response_companies}
# Count companies with no response at all
no_response_count <- sum(is.na(data$response.date)) + sum(data$response.date == "", na.rm = TRUE)
no_response_count
```
```{r no_response_companies_percent}
percent(no_response_count/total_applications)
```

```{r pie_0}

# Group the data into three categories: Yes, No, and No Response
response_data <- data %>%
  mutate(response_status = case_when(
    `selected.for.interview.` == "Yes" ~ "Interview",
    `selected.for.interview.` == "Other" ~ "No Interview",
    `selected.for.interview.` == "No" ~ "No Interview",
    `selected.for.interview.` == "TBD" ~ "No Interview",
    `selected.for.interview.` == "N/A" ~ "No Interview",
    `selected.for.interview.` == "" ~ "No Response"
  )) %>%
  group_by(response_status) %>%
  summarise(count = n())

# Create the pie chart
p3 <- ggplot(response_data, aes(x = "", y = count, fill = response_status)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +  # Turn the bar chart into a pie chart
  labs(title = "Application Responses") +
  theme_void() +  # Remove the background and grid lines for a cleaner look
  theme(legend.title = element_blank()) +  # Remove legend title
  scale_fill_manual(values = c("No Response" = my_colors[2], "Interview" = my_colors[4], "No Interview" = my_colors[7])) + 
  geom_text(aes(label = paste0(round(count / sum(count) * 100), "%")),
            position = position_stack(vjust = 0.5))  # Add percentage labels to each slice

p3

ggsave(p3, 
       filename = "response_pie.png",
       device = "png",
       height = 4, width = 4, units = "in")

```

## 5. Average Time to Interview Request

```{r avg_time_to_interview}
# Calculate the time to respond, handling NA values in response.date
data <- data %>%
  mutate(time.to.respond = as.numeric(difftime(response.date, applied.date, units = "days")))

# Calculate average time to interview request (for those who received an interview)
avg_time_to_interview_yes <- data %>%
  filter(!is.na(time.to.respond), `selected.for.interview.` == "Yes") %>%
  summarise(avg_time = mean(`time.to.respond`, na.rm = TRUE)) %>%
  pull(avg_time)  # Extract as a single value

# Calculate average time to response (for those who did not receive an interview)
avg_time_to_interview_no <- data %>%
  filter(!is.na(time.to.respond), `selected.for.interview.` == "No") %>%
  summarise(avg_time = mean(`time.to.respond`, na.rm = TRUE)) %>%
  pull(avg_time)  # Extract as a single value

# Create a table with the labeled results
average_times_table <- tibble(
  Interview_Status = c("Interview", "No Interview"),
  Avg_Time_To_Respond = c(round(avg_time_to_interview_yes, 1), round(avg_time_to_interview_no, 1))
)

# Display the table using kable
average_times_table %>%
  kable(col.names = c("Interview Status", "Average Time to Respond (Days)"),
        caption = "Average Time to Respond")

```

## 6. Job Title Analysis

```{r wordcloud}

# Preprocess job titles and create a new cleaned column in the original dataframe
data <- data %>%
  mutate(job.titles.cleaned = tolower(job.title),  # Convert to lowercase
         job.titles.cleaned = str_trim(job.titles.cleaned),  # Remove leading and trailing whitespace
         
         # Remove specific words
         job.titles.cleaned = gsub("\\bsenior\\b", "", job.titles.cleaned),  # Remove 'senior'
         job.titles.cleaned = gsub("\\bsr.\\b", "", job.titles.cleaned),  # Remove 'sr.'
         job.titles.cleaned = gsub("\\bintern\\b", "", job.titles.cleaned),  # Remove 'intern'
         job.titles.cleaned = gsub("\\btechnical\\b", "", job.titles.cleaned),  # Remove 'technical'
         job.titles.cleaned = gsub("\\bfinancial\\b", "", job.titles.cleaned),  # Remove 'financial'
         job.titles.cleaned = gsub("\\bprinciple\\b", "", job.titles.cleaned),  # Remove 'principle'
         job.titles.cleaned = gsub("\\blead\\b", "", job.titles.cleaned),  # Remove 'lead'
         job.titles.cleaned = gsub("\\bjunior\\b", "", job.titles.cleaned),  # Remove 'junior'
         job.titles.cleaned = gsub("\\bassociate\\b", "", job.titles.cleaned),  # Remove 'associate'
         job.titles.cleaned = gsub("\\bcontract\\b", "", job.titles.cleaned),  # Remove 'contract'
         job.titles.cleaned = gsub("\\bentry level\\b", "", job.titles.cleaned),  # Remove 'entry level'

         # Remove punctuation
         job.titles.cleaned = gsub("[[:punct:]]+", " ", job.titles.cleaned),  # Remove punctuation
        
         # Remove extra spaces created after word removal
         job.titles.cleaned = str_squish(job.titles.cleaned))

# Create a frequency table and filter out the specific job title "General Interest Application"
title_freq <- table(data$job.titles.cleaned[!grepl("general interest application", data$job.titles.cleaned, ignore.case = TRUE)])

# Convert the table to a data frame for sorting
title_freq_df <- as.data.frame(title_freq)

# Rename the columns for clarity
colnames(title_freq_df) <- c("Job_Title", "Frequency")

# Order the table by frequency in descending order
title_freq_df <- title_freq_df %>%
  arrange(desc(Frequency))

# View the top 10 most frequent job titles
top_title_freq <- head(title_freq_df, 10)

# Create the word cloud from the dataframe
wordcloud(words = top_title_freq$Job_Title,    # Words from the 'Job_Title' column
          freq = top_title_freq$Frequency,     # Frequencies from the 'Frequency' column
          min.freq = 3, random.order = TRUE, random.color = TRUE ,
          colors = my_colors, 
          scale = c(3.25, 1))

#top_title_freq
```


```{r job_title_distribution}
# Group job titles into broader categories
data <- data %>%
  mutate(job_category = case_when(
    grepl("data.*analyst|data analytics|data analysis", job.titles.cleaned, ignore.case = TRUE) ~ "Data Analyst",  # Matches "data" followed by any words and then "analyst"
    
    grepl("data.*engineer", job.titles.cleaned, ignore.case = TRUE) ~ "Data Engineer",  # Matches "data" followed by any words and then "engineer"
    
    grepl("data.*scientist|data science", 
          job.titles.cleaned, ignore.case = TRUE) ~ "Data Scientist",  # Flexible match for "data scientist"
    
    grepl("analytics.*engineer|business intelligence engineer|bi engineer|.*developer", 
          job.titles.cleaned, ignore.case = TRUE) ~ "Analytics Engineer",  # Flexible match for Analytics or BI roles
    
    grepl("business.*analyst|bi.*analyst|BI Consultant", 
          job.titles.cleaned, ignore.case = TRUE) ~ "Business Analyst",  # Flexible match for Business and Systems Analyst
    
    TRUE ~ "Other"
  ))



# Group by job category and summarize the count
job_distribution <- data %>%
  group_by(job_category) %>%
  summarise(count = n()) %>%
  arrange(desc(count))  # Order the categories by count, descending




# Customize the tooltip using `aes(text)` in the plot
p4 <- ggplot(job_distribution, aes(x = reorder(job_category, -count), 
                                   y = count, 
                                   fill = job_category, 
                                   text = paste("Job Category: ", job_category, 
                                                "<br>Number of Applications: ", count))) +
  
  # Create the bar plot
  geom_bar(stat = "identity", show.legend = FALSE) +  # Remove the legend
  
  # Add totals above the bars
  geom_text(aes(label = count), vjust = -0.5, size = 4) +  
  
  # Apply the chosen color palette
  scale_fill_manual(values = my_colors) +  
  
  # Customize the labels and theme
  labs(title = "Distribution of Applications by Job Title", 
       x = "Job Category", 
       y = "Number of Applications") +
  
  # Remove grey background and customize theme
  theme_minimal() +
  theme(
    legend.title= element_blank(),
    panel.grid.major = element_blank(),  # Remove major grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines
    axis.line = element_line(color = "black"),  # Add axis lines
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)  # Center the title
  ) +
  guides(fill = "none")

ggsave(p4, 
       filename = "applications_by_title.png",
       device = "png",
       height = 4, width = 6, units = "in")

# Convert to plotly and use the custom `text` for the tooltip
ggplotly(p4, tooltip = "text")

#api_create(ggplotly(p4, tooltip = "text"), filename = "Applications by Job Title")
```



```{r check}

# Select specific columns and filter for rows where job.category is 'other'
filtered_data <- data %>%
  select(job.title, job.titles.cleaned, job_category) %>%
  filter(job_category == 'Other')

# Sort the filtered data by job.titles.cleaned
filtered_data <- filtered_data %>%
  arrange(job.titles.cleaned)

# View the filtered data
#head(filtered_data)
```

```{r pie}

# Filter for applications that resulted in an interview and calculate proportions
interview_data <- data %>%
  filter(`selected.for.interview.` == "Yes") %>%
  group_by(job_category) %>%
  summarise(count = n()) %>%
  mutate(fraction = count / sum(count),
         csum = rev(cumsum(rev(fraction))),  # Cumulative sum for positioning
         pos = fraction / 2 + lead(csum, 1),  # Position labels at the middle of the segment
         pos = if_else(is.na(pos), fraction / 2, pos),  # Handle NA for the last position
          label = paste0(round(fraction * 100), "% ", job_category))  # Create percentage labels

# Create the pie chart with lines pointing to the segments
ggplot(interview_data, aes(x = "", y = fraction, fill = job_category)) +
  geom_col(width = 1, color = "white") +  # Create the pie chart with segments
  coord_polar(theta = "y") +  # Convert to a pie chart
  theme_void() +  # Remove background and axes
  labs(title = "Interviews by Job Category") +
  
  # Add percentage labels outside the pie chart with lines pointing to the segments
  geom_label_repel(aes(y = pos, label = label),
                   size = 4.5, nudge_x = 1, show.legend = FALSE, box.padding = 0.5) +

  scale_fill_manual(values = my_colors) +
  
  # Remove the legend for job categories
  guides(fill = "none")


  
```

normalized data
```{r normal_data}
# Calculate total applications and total interviews by job category
normalized_interview_data <- data %>%
  group_by(job_category) %>%
  summarise(
    total_applications = n(),  # Count total applications in each category
    total_interviews = sum(`selected.for.interview.` == "Yes", na.rm = TRUE)  # Count total interviews in each category
  ) %>%
  mutate(
    success_rate = total_interviews / total_applications  # Calculate success rate
  ) %>%
  filter(total_interviews > 0)  # Keep only categories with interviews

# Calculate fractions for labeling and positioning
normalized_interview_data <- normalized_interview_data %>%
  mutate(fraction = success_rate / sum(success_rate),  # Normalize success rates
         csum = rev(cumsum(rev(fraction))),  # Cumulative sum for positioning
         pos = fraction / 2 + lead(csum, 1),  # Position labels at the middle of the segment
         pos = if_else(is.na(pos), fraction / 2, pos),  # Handle NA for the last position
         label = paste0(round(success_rate * 100, 1), "% ", job_category))  # Create percentage labels

# Customize the tooltip using `aes(text)` in the plot
p5 <- ggplot(normalized_interview_data, aes(x = reorder(job_category, success_rate), 
                                            y = success_rate, 
                                            fill = job_category, 
                                            text = paste("Job Category: ", job_category, 
                                                         "<br>Success Rate: ", round(success_rate * 100, 1), "%"))) +
  
  # Create the bar plot with success rate
  geom_bar(stat = "identity", color = "white", show.legend = FALSE) +  # Bars with success rate
  
  # Add percentage labels above the bars
  geom_text(aes(label = paste0(round(success_rate * 100, 1), "%")), vjust = 1, size = 4) +  
  
  # Customize the labels and theme
  labs(title = "Success Rate of Applications by Job Title",
       x = "Job Category",
       y = "Success Rate (Interviews / Applications)") +
  
  # Format y-axis as percentages
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +  
  
  # Use custom color palette
  scale_fill_manual(values = my_colors) +  
  
  # Remove grey background and customize theme
  theme_minimal() +
  theme(legend.title = element_blank(),
        panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank(),  # Remove minor grid lines
        axis.line = element_line(color = "black")) +  # Add axis lines
  
  guides(fill = "none")

ggsave(p5, 
       filename = "success_rate_by_job_title.png",
       device = "png",
       height = 4, width = 6, units = "in")

# Convert to plotly and use the custom `text` for the tooltip
ggplotly(p5, tooltip = "text")


#api_create(ggplotly(p5, tooltip = "text"), filename = "Success rate by Job Title")

```


```{r bubble}
# Create a vector that maps job titles to self-assessment scores
self_assessment_values <- c("Data Scientist" = 5, 
                            "Data Analyst" = 9, 
                            "Analytics Engineer" = 9, 
                            "Data Engineer" = 6, 
                            "Business Analyst" = 9)

# Add self-assessment ratings to the normalized_interview_data dataframe
normalized_interview_data <- normalized_interview_data %>%
  mutate(self_assessment = self_assessment_values[job_category])


# Customize the tooltip using `aes(text)` in the plot
p6 <- ggplot(normalized_interview_data, aes(x = reorder(job_category, success_rate), 
                                            y = success_rate, 
                                            size = self_assessment, 
                                            color = job_category, 
                                            text = paste("Job Category: ", job_category,
                                                         "<br>Success Rate: ", round(success_rate * 100, 1), "%",
                                                         "<br>Self-Assessment (1-10): ", self_assessment))) +
  
  # Create the bubble plot
  geom_point(alpha = 1) +
  
  # Customize the size and color
  scale_color_manual(values = my_colors, guide = "none") +  # Apply custom colors but hide the color legend
  scale_size_continuous(range = c(5, 15)) +  # Set bubble sizes
  
  # Set y-axis to percent with limits from 0 to 10%
  scale_y_continuous(labels = scales::percent_format(scale = 100), limits = c(0, .10)) +  
  
  # Use a minimal theme and adjust labels
  theme_minimal() +  
  labs(title = "Success Rate of Applications by Job Category and Self-Assessment",
       x = "Job Category",
       y = "Success Rate (Interviews / Applications)",
       caption = "*Bubble size corresponds to self-assessment of skills.") +
  
  # Customize the caption and remove the legend
  theme(legend.position = "none",
        plot.caption = element_text(hjust = 0.1),  # Align caption
        plot.caption.position = "plot") 

# Convert to plotly and use the custom `text` for the tooltip
ggplotly(p6, tooltip = "text")

ggsave(p6, 
       filename = "self_assessment.png",
       device = "png",
       height = 4, width = 6, units = "in")

#api_create(ggplotly(p6, tooltip = "text"), filename = "Success rate by Job Title and Self Assessment")
```

## 7. Job Post Age


average age of posts when I applied
```{r age}
  # Calculate job.post.age
data <- data %>%
  mutate(job.post.age = as.numeric(difftime(`applied.date`, `job.post.date`, units = "days")))

# Function to remove outliers based on IQR
remove_outliers <- function(x) {
  Q1 <- quantile(x, 0.25, na.rm = TRUE)  # First quartile (25th percentile)
  Q3 <- quantile(x, 0.75, na.rm = TRUE)  # Third quartile (75th percentile)
  IQR <- Q3 - Q1  # Interquartile range
  
  # Define outlier boundaries
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  
  # Filter values that are within the lower and upper bound
  x[x >= lower_bound & x <= upper_bound]
}

# Apply the function to remove outliers and calculate the mean
cleaned_data <- remove_outliers(data$job.post.age)

# Calculate mean without outliers
mean_without_outliers <- mean(cleaned_data, na.rm = TRUE)

# Print the average age
mean_without_outliers

```


```{r hist}

# 1. Calculate the percentage of jobs with job.post.age <= 2
total_jobs <- nrow(data)
jobs_within_2_days <- sum(cleaned_data <= 2, na.rm = TRUE)
percentage_within_2_days <- (jobs_within_2_days / total_jobs) * 100

# Print the percentage
cat("Percentage of jobs applied to within 2 days of posting:", round(percentage_within_2_days, 2), "%\n")

# 2. Calculate the standard deviation of 'job.post.age'
mean_days <- mean(cleaned_data, na.rm = TRUE)  # Calculate the mean
sd_days <- sd(cleaned_data, na.rm = TRUE)  # Calculate the standard deviation

# Print the mean and standard deviation
cat("Mean days since job posted:", round(mean_days, 2), "\n")
cat("Standard deviation:", round(sd_days, 2), "\n")

# 3. One standard deviation from the mean
lower_bound <- mean_days - sd_days
upper_bound <- mean_days + sd_days

cat("One standard deviation from the mean is between", round(lower_bound, 2), "and", round(upper_bound, 2), "days.\n")

```

Job post age and interview success
```{r}
# Compare average job post age for successful vs. unsuccessful applications
# Filter the data to include only "Yes" or "No" for selected.for.interview., ignoring everything else and NA
age_success_data <- data %>%
 # filter(`selected.for.interview.` %in% c("Yes", "No")) %>%  # Include only "Yes" or "No" responses
  group_by(`selected.for.interview.`) %>%
  summarise(average_job_post_age = mean(job.post.age, na.rm = TRUE))

# Create a bar plot to visualize the connection between interview success and job post age
ggplot(age_success_data, aes(x = `selected.for.interview.`, y = average_job_post_age, fill = `selected.for.interview.`)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  geom_text(aes(label = round(average_job_post_age, 2)), vjust = -0.5, size = 3.5) +
  labs(title = "Average Age of Job Posts by Interview Success", x = "Interview Success (Yes/No)", y = "Average Job Post Age (Days)") +
  theme_minimal() +
scale_fill_manual(values = my_colors)

```

## 8. Job Source Evaluation

Job source effectiveness

```{r job_source}
# Group job sources where applications resulted in an interview, group the rest as 'Other', and exclude 'Recruiter Outreach'
source_data <- data %>%
  # Exclude "Recruiter Outreach"
  filter(Job.Source != "Recruiter Outreach") %>%
  
  group_by(Job.Source) %>%
  summarise(total_applications = n(),
            total_interviews = sum(`selected.for.interview.` == "Yes", na.rm = TRUE)) %>%
  
  # Only keep job sources that resulted in at least one interview, group others as 'Other'
  mutate(Job.Source = if_else(total_interviews > 0, Job.Source, "Other")) %>%
  
  # Group by the updated Job.Source and calculate totals
  group_by(Job.Source) %>%
  summarise(total_applications = sum(total_applications),
            total_interviews = sum(total_interviews)) %>%
  
  # Order by the total number of applications (descending)
  arrange(desc(total_applications))

# Customize the tooltip using `aes(text)` in `geom_bar`
p7 <- ggplot(source_data, aes(x = reorder(Job.Source, -total_applications), 
                              y = total_applications, 
                              fill = Job.Source, 
                              text = paste("Job Source: ", Job.Source, 
                                           "<br>Total Applications: ", total_applications))) +
  
  # Bar plot
  geom_bar(stat = "identity", show.legend = FALSE) +
  
  # Add total number of applications as labels
  geom_text(aes(label = total_applications), 
            vjust = 1, size = 3.5) +
  
  # Customize the labels and theme
  labs(title = "Applications by Job Source", 
       x = "Job Source", 
       y = "Total Applications") +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  scale_fill_manual(values = my_colors) +
  guides(fill = "none")

ggsave(p7, 
       filename = "job_boards.png",
       device = "png",
       height = 4, width = 6, units = "in")

# Convert to plotly and customize the tooltip display to use `text`
ggplotly(p7, tooltip = "text")

#api_create(ggplotly(p7, tooltip = "text"), filename = "Applications by Job Source")
```

```{r stacked_source_outcomes}
# Calculate success rate for each job source
source_data <- source_data %>%
  mutate(success_rate = (total_interviews / total_applications) * 100)

# Customize the tooltip using `aes(text)` in `geom_bar`
p8 <- ggplot(source_data, aes(x = reorder(Job.Source, -success_rate), 
                              y = success_rate, 
                              fill = Job.Source, 
                              text = paste("Job Source: ", Job.Source, 
                                           "<br>Success Rate: ", round(success_rate, 1), "%"))) +
  
  # Bar plot with white borders around bars
  geom_bar(stat = "identity", color = "white") +
  
  # Add percentage success rate as labels
  geom_text(aes(label = paste0(round(success_rate, 1), "%")),
            vjust = 1, size = 3.5) +

  # Customize labels and theme
  labs(title = "Success Rate by Job Source", 
       x = "Job Source", 
       y = "Success Rate (%)") +
  theme_minimal() +
  theme(legend.title= element_blank()) +
  scale_fill_manual(values = my_colors) +
  guides(fill = "none")

ggsave(p8, 
       filename = "success_by_job_board.png",
       device = "png",
       height = 4, width = 6, units = "in")

# Convert to plotly and customize the tooltip display to use `text`
ggplotly(p8, tooltip = "text")

#api_create(ggplotly(p8, tooltip = "text"), filename = "Success Rate by Job Source")

```

